---
title: "Data manipulation: Maternal diet quality and brain morphology"
author: "Yuchan Mou"
date: "16/Mar/2022"
output: html_document
---

# Set up in R studio 
```{r global settings, include=FALSE, message=FALSE}
# Global settings
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.table.format = "html")
```

```{r set working directory, include=FALSE, message=FALSE}
.libPaths(r"(V:\HomeDir\044073(J. Mou)\Projects\temp_wd_r\library)")

# WD settings
getwd()

load(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\DataManipulation_maternaldietBrain_2022-11-30.RData)')
```

````{r installpck, message = FALSE}
packages <- c('haven', "tidyverse")
invisible(lapply(packages, library, character.only = T))

systime <- Sys.Date()
analysis <- "maternaldietBrain"
```


# Load data 

```{r exposure}
mdietq <- haven::read_spss(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\MOTHERDietScore_pregnancy_18052018.sav)') %>%
  select('IDM', 'Energy_kcal_pregnancy', 'DietScore_pregnancy') %>% 
  arrange(IDM)
# individual components are not included in the dataset. Need to ask Trudy
```


```{r outcomes}
## general MRI info data
bcore <- readRDS(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\genr_mri_core_data_20220311.rds)') %>% 
  select(idc,age_child_mri_f09, mri_consent_f09, t1_has_nii_f09, t1_asset_has_nii_f09, has_braces_mri_f09, exclude_incidental_f09, freesurfer_qc_f09, dti_has_nii_f09, dti_man_qc_f09, dti_auto_qc_f09, dti_overall_qc_f09,
         qdec_has_lgi_f09,
         age_child_mri_f13, mri_consent_f13, t1_has_nii_f13, has_braces_mri_f13, exclude_incidental_f13, freesurfer_qc_f13, dti_has_nii_f13, dti_man_qc_f13, dti_auto_qc_f13, dti_overall_qc_f13, qdec_has_lgi_f13) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

path <- c(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\F09\)',
          r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\F13\)')

#------------------------  
# F09
## total brain volume, cerebral white/gray matter, subcortical gray matter volumes
tbv_09 <- readRDS(paste(path[1], r'(f09_freesurfer_v6_09dec2016_tbv_stats_pull20june2017_v2.rds)', sep = "")) %>% 
  select(idc, CerebralWhiteMatterVol_f09, CortexVol_f09, TotalGrayVol_f09, SubCortGrayVol_f09, SupraTentorialVol_f09, genr_tbv_f09, eTIV_f09) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

## prefrontal cortical volume
pfc_09_1 <- readRDS(paste(path[1], r"(f09_freesurfer_v6_09dec2016_lobes_stats_pull06june2017.rds)", sep = "")) %>% 
  select(idc, lh_frontal_vol_f09, rh_frontal_vol_f09) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

pfc_09_2 <- readRDS(paste(path[1], r"(f09_freesurfer_v6_09dec2016_aparc_stats_pull06june2017.rds)", sep = "")) %>% 
  select(idc, lh_paracentral_vol_f09, rh_paracentral_vol_f09, lh_precentral_vol_f09, rh_precentral_vol_f09) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

## hippocampus and amygdala
seg_09 <- readRDS(paste(path[1], r"(f09_freesurfer_v6_09dec2016_aseg_stats_pull06june2017_v1.rds)", sep = "")) %>% 
  select(idc, Left_Amygdala_vol_f09, Right_Amygdala_vol_f09, Left_Hippocampus_vol_f09, Right_Hippocampus_vol_f09) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

## Fractional anisotropy and Mean diffusivity
dti_f09 <- readRDS(paste(path[1], r"(f09_GenR_MRI_eddy_dipy_wls_14Feb2022_autoPtx_dti_stats_v1.rds)", sep = "")) %>% 
  select(idc, cst_l_dti_dipy_wls_mean_FA_f09, cst_r_dti_dipy_wls_mean_FA_f09, 
         cgc_l_dti_dipy_wls_mean_FA_f09, cgc_r_dti_dipy_wls_mean_FA_f09,
         cgh_l_dti_dipy_wls_mean_FA_f09, cgh_r_dti_dipy_wls_mean_FA_f09, 
         unc_l_dti_dipy_wls_mean_FA_f09, unc_r_dti_dipy_wls_mean_FA_f09, 
         cst_l_dti_dipy_wls_mean_MD_f09, cst_r_dti_dipy_wls_mean_MD_f09, 
         cgc_l_dti_dipy_wls_mean_MD_f09, cgc_r_dti_dipy_wls_mean_MD_f09, 
         cgh_l_dti_dipy_wls_mean_MD_f09, cgh_r_dti_dipy_wls_mean_MD_f09, 
         unc_l_dti_dipy_wls_mean_MD_f09, unc_r_dti_dipy_wls_mean_MD_f09) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)


#------------------------  
# F13
## total brain volume, cerebral white/gray matter, subcortical gray matter volumes
tbv_13 <- readRDS(paste(path[2], r'(f13_freesurfer_v6_14oct2020_tbv_stats_pull23Nov2020_v2.rds)', sep = "")) %>% 
  select(idc, CerebralWhiteMatterVol_f13, CortexVol_f13, TotalGrayVol_f13, SubCortGrayVol_f13, SupraTentorialVol_f13, eTIV_f13, genr_tbv_f13) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

## prefrontal cortical volume
pfc_13_1 <- readRDS(paste(path[2], r"(f13_freesurfer_v6_29may2021_lobes_stats_pull23Nov2020.rds)", sep = "")) %>% 
  select(idc, lh_frontal_vol_f13, rh_frontal_vol_f13) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

pfc_13_2 <- readRDS(paste(path[2], r"(f13_freesurfer_v6_14oct2020_aparc_stats_pull23Nov2020.rds)", sep = "")) %>% 
  select(idc, lh_paracentral_vol_f13, rh_paracentral_vol_f13, lh_precentral_vol_f13, rh_precentral_vol_f13) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

## hippocampus and amygdala
seg_13 <- readRDS(paste(path[2], r"(f13_freesurfer_v6_14oct2020_aseg_stats_pull23Nov2020_v1.rds)", sep = "")) %>% 
  select(idc, Left_Amygdala_vol_f13, Right_Amygdala_vol_f13, Left_Hippocampus_vol_f13, Right_Hippocampus_vol_f13) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)


## Fractional anisotropy and Mean diffusivity
dti_f13 <- readRDS(paste(path[2], r"(f13_GenR_MRI_eddy_dipy_wls_14Feb2022_autoPtx_dti_stats_v1.rds)", sep = "")) %>% 
  select(idc, cst_l_dti_dipy_wls_mean_FA_f13, cst_r_dti_dipy_wls_mean_FA_f13, 
         cgc_l_dti_dipy_wls_mean_FA_f13, cgc_r_dti_dipy_wls_mean_FA_f13,
         cgh_l_dti_dipy_wls_mean_FA_f13, cgh_r_dti_dipy_wls_mean_FA_f13, 
         unc_l_dti_dipy_wls_mean_FA_f13, unc_r_dti_dipy_wls_mean_FA_f13, 
         cst_l_dti_dipy_wls_mean_MD_f13, cst_r_dti_dipy_wls_mean_MD_f13, 
         cgc_l_dti_dipy_wls_mean_MD_f13, cgc_r_dti_dipy_wls_mean_MD_f13, 
         cgh_l_dti_dipy_wls_mean_MD_f13, cgh_r_dti_dipy_wls_mean_MD_f13, 
         unc_l_dti_dipy_wls_mean_MD_f13, unc_r_dti_dipy_wls_mean_MD_f13) %>% 
  rename(IDC = idc) %>% 
  arrange(IDC)

```

```{r cov}
#all covariates
cov <- haven::read_spss(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\CHILD-ALLGENERALDATA_12112020.sav)') %>%  
  select('IDC','IDM', "MOTHER", "PARTNER", "MULTIPLE", "RANKCHILD", 'TWIN', 'GESTBIR', 'WEIGHT',# twin birth, gestational age at birth, weight live birth
         "COHORT1", "COHORT2", "FUPFASE1", "FASE1", "FASE2","FUPFASE2", "FUPFASE2_END", "STARTFASE3_5", "FUPFASE3_5", 
         "VISIT5", "STARTFASE3_9", "FUPFASE3_9", "VISIT9", "FUPFASE4_13", "VISIT13", "INTAKE",
         'AGE_M_v2', 'BMI_0', "PARITY", # parity
         'GENDER',  'EDUCM', 'EDUCVM','INCOME', 'INCOME5', 'ETHNMv2', "ETHNINFv2") %>% 
  mutate_at(c("GENDER", "EDUCM", "EDUCVM", 
         "INCOME5", "INCOME", 
         'ETHNINFv2', "ETHNMv2"), as_factor) %>% 
  arrange(IDM)


## smoking
cov_smoke <- haven::read_spss(r'(V:\HomeDir\044073(J. Mou)\Projects\3 Dietary patterns and brain morphology\2 data\cov\MATERNALSMOKING_22112016.sav)') %>%
  select('idm', 'SMOKE_ALL') %>% 
  rename(IDM = idm)

## alcohol
cov_alcoh <- haven::read_spss(r'(V:\HomeDir\044073(J. Mou)\Projects\3 Dietary patterns and brain morphology\2 data\cov\GEDRAGSGROEP_MaternalDrinking_22112016.sav)') %>% 
  select('IDM', 'mdrink_updated') %>% 
  mutate_at(c('mdrink_updated'), as_factor) %>% 
  arrange(IDM)

## folic acid
cov_folic <- haven::read_spss(r'(V:\HomeDir\044073(J. Mou)\Projects\3 Dietary patterns and brain morphology\2 data\cov\MATERNALFOLICACID_23062010.sav)') %>% 
  mutate_at(c('FOLIUM_VALIDATED'), as_factor) %>% 
  arrange(IDM)
  
## BSI-prenatal
cov_bsi <- haven::read_spss(r'(V:\HomeDir\044073(J. Mou)\Projects\3 Dietary patterns and brain morphology\2 data\cov\GR1003-BSI D1_22112016.sav)') %>% 
  select('IDM', 'gsi') %>% 
  arrange(IDM)

## pregnancy complication
## preeclampsie is validated. Diabetes, preexisting hypertension and pregnancy induced hypertension is not validated. 
cov_pregcomp <- haven::read_spss(r"(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\MATERNALCOMPLICATIONS_22112016.sav)") %>% 
  arrange(IDM)

## puberty stage
# can't find dataset. waiting for Lorenza's reply

## breastfeeding 
cov_breastfeeding <- haven::read_spss(r"(V:\HomeDir\044073(J. Mou)\Projects\2 Meta analysis - Obesity review-GenR\9. prediction model\data\CHILDBREASTFEEDING_18122017.sav)") %>% 
  arrange(IDC) %>% 
  select(IDC, AGESTOPB)

## child diet quality 8y - mediator
cov_cdq <- haven::read_spss(r"(V:\HomeDir\044073(J. Mou)\Projects\3 Dietary patterns and brain morphology\2 data\diet\CHILDDietScore_8y_18052018.sav)") %>% 
  select("IDC", 'DietScore_8y') %>% 
  arrange(IDC)

## child IQ
fsiq_13y <- haven::read_spss(r"(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\CHILDWISC13_16082021.sav)") %>% 
  arrange(IDC) %>% 
  select(IDC, visitF13, AGECHILD13, WISC13_FSIQ, WISC13_Voc_Tscore, WISC13_MR_Tscore, WISC13_DS_Tscore, WISC13_CD_Tscore)

```

```{r merge dataset}
dat_m <- mdietq %>% 
  left_join(cov, by = 'IDM') %>% 
  left_join(cov_smoke, by = 'IDM') %>% 
  left_join(cov_alcoh, by = 'IDM') %>% 
  left_join(cov_folic, by = 'IDM') %>% 
  left_join(cov_bsi, by = 'IDM') %>%
  left_join(cov_pregcomp, by = "IDM") %>% 
  arrange(IDC)

dat_c <- bcore %>% 
  left_join(tbv_09, by = 'IDC') %>% 
  left_join(tbv_13, by = 'IDC') %>% 
  left_join(pfc_09_1, by = "IDC") %>% 
  left_join(pfc_09_2, by = 'IDC') %>%
  left_join(pfc_13_1, by = "IDC") %>% 
  left_join(pfc_13_2, by = 'IDC') %>%
  left_join(seg_09, by = 'IDC') %>% 
  left_join(seg_13, by = "IDC") %>% 
  left_join(dti_f09, by = "IDC") %>% 
  left_join(dti_f13, by = "IDC") %>% 
  left_join(cov_cdq, by = "IDC") %>% 
  left_join(cov_breastfeeding, by = "IDC") %>% 
  left_join(fsiq_13y, by = "IDC") %>% 
  arrange(IDC)

dat <- dat_c %>% 
  arrange(IDC) %>% 
  left_join(dat_m, by = 'IDC') %>% 
  filter(!is.na("IDC"))
```


# Data manipulation 

## 1.Recode variables

### 1.1 volumes calculation and unit conversion
```{r volumes and unit calculation}
dat <- dat %>% mutate(#9y
               # pfc mm3 -> cm3
               lh_prefrontal_vol_f09 = lh_frontal_vol_f09 - lh_paracentral_vol_f09 - lh_precentral_vol_f09,
               rh_prefrontal_vol_f09 = rh_frontal_vol_f09 - rh_paracentral_vol_f09 - rh_precentral_vol_f09,
               prefrontal_vol_f09 = lh_prefrontal_vol_f09 + rh_prefrontal_vol_f09,
               amygdala_vol_f09 = Left_Amygdala_vol_f09 + Right_Amygdala_vol_f09,
               hippocampus_vol_f09 = Left_Hippocampus_vol_f09 + Right_Hippocampus_vol_f09, 
               #FA
               glb_mean_FA_f09 = (cst_l_dti_dipy_wls_mean_FA_f09 + cst_r_dti_dipy_wls_mean_FA_f09)/2,
               cigulum_mean_FA_f09 = (cgc_l_dti_dipy_wls_mean_FA_f09 + cgc_r_dti_dipy_wls_mean_FA_f09 + 
                                      cgh_l_dti_dipy_wls_mean_FA_f09 + cgh_r_dti_dipy_wls_mean_FA_f09)/4,
               uncinate_mean_FA_f09 = (unc_l_dti_dipy_wls_mean_FA_f09 + unc_r_dti_dipy_wls_mean_FA_f09)/2,
               # MD
               glb_mean_MD_f09 = (cst_l_dti_dipy_wls_mean_MD_f09 + cst_r_dti_dipy_wls_mean_MD_f09)/2,
               cigulum_mean_MD_f09 = (cgc_l_dti_dipy_wls_mean_MD_f09 + cgc_r_dti_dipy_wls_mean_MD_f09 + 
                                      cgh_l_dti_dipy_wls_mean_MD_f09 + cgh_r_dti_dipy_wls_mean_MD_f09)/4,
               uncinate_mean_MD_f09 = (unc_l_dti_dipy_wls_mean_MD_f09 + unc_r_dti_dipy_wls_mean_MD_f09)/2,
               
               # 13y
               # pfc mm3 -> cm3
               lh_prefrontal_vol_f13 = lh_frontal_vol_f13 - lh_paracentral_vol_f13 - lh_precentral_vol_f13,
               rh_prefrontal_vol_f13 = rh_frontal_vol_f13 - rh_paracentral_vol_f13 - rh_precentral_vol_f13,
               prefrontal_vol_f13 = lh_prefrontal_vol_f13 + rh_prefrontal_vol_f13,
               amygdala_vol_f13 = Left_Amygdala_vol_f13 + Right_Amygdala_vol_f13,
               hippocampus_vol_f13 = Left_Hippocampus_vol_f13 + Right_Hippocampus_vol_f13,
               # FA
               glb_mean_FA_f13 = (cst_l_dti_dipy_wls_mean_FA_f13 + cst_r_dti_dipy_wls_mean_FA_f13)/2,
               cigulum_mean_FA_f13 = (cgc_l_dti_dipy_wls_mean_FA_f13 + cgc_r_dti_dipy_wls_mean_FA_f13 + 
                                      cgh_l_dti_dipy_wls_mean_FA_f13 + cgh_r_dti_dipy_wls_mean_FA_f13)/4,
               uncinate_mean_FA_f13 = (unc_l_dti_dipy_wls_mean_FA_f13 + unc_r_dti_dipy_wls_mean_FA_f13)/2,
               # MD
               glb_mean_MD_f13 = (cst_l_dti_dipy_wls_mean_MD_f13 + cst_r_dti_dipy_wls_mean_MD_f13)/2,
               cigulum_mean_MD_f13 = (cgc_l_dti_dipy_wls_mean_MD_f13 + cgc_r_dti_dipy_wls_mean_MD_f13 + 
                                      cgh_l_dti_dipy_wls_mean_MD_f13 + cgh_r_dti_dipy_wls_mean_MD_f13)/4,
               uncinate_mean_MD_f13 = (unc_l_dti_dipy_wls_mean_MD_f13 + unc_r_dti_dipy_wls_mean_MD_f13)/2
               ) %>% 
  mutate(# unit conversion mm3 -> cm3
        across(c(prefrontal_vol_f09, CerebralWhiteMatterVol_f09, CortexVol_f09, TotalGrayVol_f09, SubCortGrayVol_f09, SupraTentorialVol_f09, genr_tbv_f09, eTIV_f09,
                 prefrontal_vol_f13, CerebralWhiteMatterVol_f13, CortexVol_f13, TotalGrayVol_f13, SubCortGrayVol_f13, SupraTentorialVol_f13, genr_tbv_f13, eTIV_f13), ~ .x/1000)
        )

```


### 1.2 maternal education
```{r}
#fct_count(dat$EDUCVM, prop = T)
# 'no education finished' and 'primary' only account for 0.05%.  Therefore collapse primary and secondary together

dat$edu <- dat$EDUCVM %>% 
  fct_collapse(high = c("higher, phase 1", "higher, phase 2"),
               low = c("secondary, phase 1", "secondary, phase 2", 'no education finished', "primary")) %>% 
  as.factor()

# auxiliary vars
dat$edu_a1 <- dat$EDUCM %>% 
  fct_collapse(high = c("higher, phase 1", "higher, phase 2"),
               low = c("secondary, phase 1", "secondary, phase 2", 'no education finished', "primary")) %>% 
  as.factor()


```

### 1.3 Household income
```{r}
fct_count(dat$INCOME)

dat$income_r <- dat$INCOME %>% 
  fct_collapse(
    '< 1200' = c("less than 450", 
                 "450-600 euro", 
                 "600-700 euro", 
                 "700-800 euro", 
                 "800-900 euro", 
                 "900-1200 euro"), 
    '1200 - 2200' = c("1200-1400 euro", 
                      "1400-1600 euro", 
                      "1600-1800 euro", 
                      "1800-2000 euro", 
                      "2000-2200 euro"), 
    '> 2200' = c('> 2200 euro')) %>% 
  as.factor()

#auxiliary variables
fct_count(dat$INCOME5)

dat$income_a1 <- dat$INCOME5 %>% 
  fct_collapse(
    '< 1200' = c("Less than € 800", 
                  "€ 800-1200"),
    '1200 - 2400' = c("€ 1200-1600",
                  "€ 1600-2000",
                  "€ 2000-2400"), 
    '> 2400' = c("€ 2400-2800",
                  "€ 2800-3200",
                  "€ 3200-4000",
                  "€ 4000-4800",
                  "€ 4800-5600",
                  "More than € 5600")) %>% 
  as.factor()

```

### 1.4 maternal ethnicity
```{r}
fct_count(dat$ETHNMv2)

dat$ethm <- dat$ETHNMv2 %>% 
  fct_collapse(
    'Dutch' = c('Dutch'), 
    'non‐Dutch Western' = c('Indonesian','American,western','Asian, western', 'European', "Oceanie"),
    'non-Dutch non-Western' = c( 'Cape Verdian', 'Moroccan', 'Dutch Antilles', 'Surinamese', 'Turkish', 'African', 'American, non western', 'Asian, non western')
    ) %>% 
  as.factor()

dat$ethc <- dat$ETHNINFv2 %>% 
  fct_collapse(
    'Dutch' = c('Dutch'), 
    'non‐Dutch Western' = c('Indonesian','American,western','Asian, western', 'European', "Oceanie"),
    'non-Dutch non-Western' = c( 'Cape Verdian', 'Moroccan', 'Dutch Antilles', 'Surinamese', 'Turkish', 'African', 'American, non western', 'Asian, non western')
    ) %>% 
  as.factor()

```

### 1.5 Mothers: alcohol during pregnancy
incl. in the diet quality score. examine in a separate model.

```{r}
fct_count(dat$mdrink_updated)

dat$drinkm <- dat$mdrink_updated %>% 
  fct_collapse(
    'mother continued drinking in pregnancy' = c(
      'mother continued drinking occasionally',
      'mother continued drinking frequently (1 or more glass/week for at least 2 trimesters)'
    )
  ) %>% 
  as.factor()
```

### 1.6 Mothers: folic acid during pregnancy
incl. in the diet quality score. examine in a separate model.

```{r}
fct_count(dat$FOLIUM_VALIDATED)

dat$folic_acid <- dat$FOLIUM_VALIDATED %>% 
  as.factor()
```

### 1.7 Mothers: Smoking during pregnancy
```{r}
dat$SMOKE_ALL <- dat$SMOKE_ALL %>%  
  as.factor()
fct_count(dat$SMOKE_ALL)
```

### 1.8 Mothers: breastfeeding duration
```{r}
dat$breastfeeding <- ifelse(dat$AGESTOPB < 2, 1, 
                            ifelse(dat$AGESTOPB >= 2 & dat$AGESTOPB < 6, 2,
                                   ifelse(dat$AGESTOPB >= 6, 3, NA)))

dat$breastfeeding <- factor(dat$breastfeeding, levels = c(1,2,3), labels = c("< 2 months", "2-5.9 months", ">= 6 months"),  ordered = F)

table(dat$breastfeeding, useNA = "always") 
```

# 1.9 Parity
```{r}
dat$parity <- ifelse(dat$PARITY == 0, 1,
                     ifelse(dat$PARITY >1, 2, NA))
dat$parity <- factor(dat$parity, levels = c(1,2), labels = c("nulliparity", "multiparity"))

table(dat$parity, useNA = "always")
```



# Study population 

```{r dataset 9y}
dat_9y <- dat %>% 
  # Mothers completed valid FFQ during pregnancy
  # n = 6485
  filter(!is.na(DietScore_pregnancy)) %>% 
  
  # Children who participated in phase 3, focus 9 years
  # -> Exclude loss to follow-up# or did not participate at 9-10 years n = 1357
  filter(FUPFASE3_9 == 1) %>% 
  # 5128 participated at 9-10 years phase
  
  # -> Exclude no consent for MRI n = 2348
  filter(mri_consent_f09 == "yes") %>% 
  # n = 2780 underwent MRI scan
  
  # -> exclude without structural MRI data n = 33
  filter(t1_has_nii_f09 == "yes"& t1_asset_has_nii_f09 != "exclude") %>% 
  # n = 2747 with structural MRI data
  
  # exclude children with braces n = 65
  filter(has_braces_mri_f09 == "no") %>%
  # n = 2682
  
  # -> Exclude no/poor reconstruction on FreeSurfer n = 421
  filter(freesurfer_qc_f09 == "usable") %>% 
  # n = 2261 with usable structural MRI data
  
  # -> Exclude major incidental findings by radiological inspection n = 12
  filter(exclude_incidental_f09 == "include") %>% 
  
  # -> exclude participants without gyrification data n = 26
  filter(qdec_has_lgi_f09 == "yes")
# FINAL n = 2223

```

```{r dataset 13y}
dat_13y <- dat %>% 
  # Mothers completed valid FFQ during pregnancy
  # n = 6485
  filter(!is.na(DietScore_pregnancy)) %>% 
  
  # Children who participated in phase 4, focus 13 years
  # -> Exclude loss to follow-up# or did not participate at 13 years n = 1663
  filter(FUPFASE4_13 == 1) %>% 
  # 4822 participated at 13 years phase
  
  # -> Exclude no consent for MRI n = 2101
  filter(mri_consent_f13 == "yes") %>% 
  # n = 2721 underwent MRI scan
  
  # -> exclude without structural MRI data n = 80
  filter(t1_has_nii_f13 == "yes") %>% 
  # n = 2641 with structural MRI data
  
  # exclude children with braces n = 764
  filter(has_braces_mri_f13 == "no") %>%
  # n = 1877
  
  # -> Exclude no/poor reconstruction on FreeSurfer n = 276
  filter(freesurfer_qc_f13 == "usable") %>% 
  # n = 1601 with usable structural MRI data
  
  # -> Exclude major incidental findings by radiological inspection n = 9
  filter(exclude_incidental_f13 == "include")  %>% 
  
  # -> Exclude participants without gyrification data n = 10
  filter(qdec_has_lgi_f13 == "yes")
# FINAL n = 1582

```

```{r longitudinal dataset}
dat_9_13y <- dat %>% 
  # Mothers completed valid FFQ during pregnancy
  # n = 6485
  filter(!is.na(DietScore_pregnancy)) %>% 
  
  # Children who participated in phase 4, focus 13 years
  # -> Exclude loss to follow-up# or did not participate at 13 years n = 1357
  filter(FUPFASE4_13 == 1) %>% 
  # 5128 participated at 13 years phase
  
  # -> Exclude no consent for MRI n = 2348
  filter(mri_consent_f09 == "yes" & mri_consent_f13 == "yes") %>% 
  # n = 2780 underwent MRI scan
  
  # -> exclude without structural MRI data n = 33
  filter(t1_has_nii_f09 == "yes"& t1_asset_has_nii_f09 != "exclude" & t1_has_nii_f13 == "yes") %>% 
  # n = 2747 with structural MRI data
  
  # exclude children with braces n = 65
  filter(has_braces_mri_f09 == "no" & has_braces_mri_f13 == "no") %>%
  # n = 2682
  
  # -> Exclude no/poor reconstruction on FreeSurfer n = 421
  filter(freesurfer_qc_f09 == "usable" & freesurfer_qc_f13 == "usable") %>% 
  # n = 2261 with usable structural MRI data
  
  # -> Exclude major incidental findings by radiological inspection n = 12
  filter(exclude_incidental_f09 == "include" & exclude_incidental_f13 == "include") %>% 
  
  # -> Exclude participants without gyrification data n = 16
  filter(qdec_has_lgi_f09 == "yes" & qdec_has_lgi_f13 == "yes")
# FINAL n = 872

```

# Save datasets 
```{r save datasets, include = FALSE}
save(dat, dat_9y, dat_13y, dat_9_13y, file = paste(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\DataManipulation_)', analysis, "_", systime, '.RData', sep = ''))

```


