---
title: 'Overview for Imputation: Maternal diet and brain morphology'
author: "Yuchan Mou"
date: "17/Mar/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global settings, include=FALSE, message=FALSE}
# Global settings

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.table.format = "html")
```

```{r set working directory, include=FALSE, message=FALSE}
# WD settings
getwd()

```

````{r installpck, message = FALSE}
# install packages
Sys.setenv(JAGS_HOME = r"(V:\HomeDir\044073(J. Mou)\program\JAGS)")

packages <- c('haven', 'tidyverse', 'sjmisc', 'ggcorrplot', 'kableExtra', 'mice', 'JointAI')
invisible(lapply(packages, library, character.only = T))
```

------
```{r load data}
load(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\DataManipulation_maternaldietBrain_2022-03-17.RData)')
```


------
## DATA 9 years
## Description: unimputed data

### 1. Distribution

The plots below shows the distributions of exposures, outcomes and covariates.

```{r,  fig.width = 10 , fig.height = 10}
dat_num <- dat_9y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy, AGE_M_v2, 
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09, 
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         age_child_mri_f09, age_child_mri_f13, DietScore_8y, gsi, BMI_0) %>% 
  as.data.frame()

par(mfrow = c(4,4))

plot_all(dat_num)

```


```{r,  fig.width = 10 , fig.height = 10}
dat_cat <- dat_9y %>% 
  select(GENDER, edu, income_r, ethm, ethc, SMOKE_ALL, drinkm, folic_acid, breastfeeding) %>% 
  as.data.frame()

par(mfrow = c(3,3))

plot_all(dat_cat)


```

### 2. Correlation matrix
```{r,  fig.width = 15 , fig.height = 15}
dat_9y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy,
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09,
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f09, 
         GENDER, edu, edu_a1, 
         income_r, income_a1,
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding) %>% 
  mutate_if(is.factor, as.numeric) %>% 
  cor(use = 'pairwise.complete.obs', method = 'spearman') %>% 
  ggcorrplot(hc.order = TRUE, 
             type = "lower",
             lab = TRUE)

```

### 3. Missing percentage
```{r}
dat_impcheck <- dat_9y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy,
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09,
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f09, 
         GENDER, edu, edu_a1, 
         income_r, income_a1,
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding) %>% 
  as.data.frame()

# Proportion of missing values 
prop_missing <- cbind(
  "#" = table(ifelse(complete.cases(dat_impcheck), 'complete', 'incompl.')),
  '%' = round(100 * sort(table(complete.cases(dat_impcheck)))/nrow(dat_impcheck) , 2)
)

# Proportion of missing values per variable
prop_missing_pervar <- as.data.frame(cbind(
  "# NA" = sort(colSums(is.na(dat_impcheck))),
  "% NA" = round(sort(colMeans(is.na(dat_impcheck))) * 100, 2)))

kable(prop_missing) %>%
  kable_styling(bootstrap_options = c("striped", "hover", 'responsive'), full_width = F )

prop_missing_pervar %>% 
  filter(`# NA` != 0) %>% 
  tibble::rownames_to_column(var = 'Variables') %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", 'condensed'), full_width = T )

# write.table(tbl_missing, file = 'proportion of missings_1y.txt', sep = ',', quote = F, row.names = F)

```


## Covariates imputation

### 1. Missing patterns

```{r missing patterns 9y,  fig.width = 10 , fig.height = 30}

md_pattern(dat_impcheck, pattern = T)$plot

```


------
## DATA 13 years
## Description: unimputed data

### 1. Distribution

The plots below shows the distributions of exposures, outcomes and covariates.

```{r,  fig.width = 10 , fig.height = 10}
dat_num <- dat_13y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy, AGE_M_v2, 
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09, 
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         age_child_mri_f09, age_child_mri_f13, DietScore_8y, gsi, BMI_0) %>% 
  as.data.frame()

par(mfrow = c(4,4))

plot_all(dat_num)

```


```{r,  fig.width = 10 , fig.height = 10}
dat_cat <- dat_13y %>% 
  select(GENDER, edu, income_r, ethm, ethc, SMOKE_ALL, drinkm, folic_acid, breastfeeding) %>% 
  as.data.frame()

par(mfrow = c(3,3))

plot_all(dat_cat)


```

### 2. Correlation matrix
```{r,  fig.width = 15 , fig.height = 15}
dat_13y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy,
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f13, 
         GENDER, edu, edu_a1, 
         income_r, income_a1,
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding) %>% 
  mutate_if(is.factor, as.numeric) %>% 
  cor(use = 'pairwise.complete.obs', method = 'spearman') %>% 
  ggcorrplot(hc.order = TRUE, 
             type = "lower",
             lab = TRUE)

```

### 3. Missing percentage
```{r}
dat_impcheck <- dat_13y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy,
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f13, 
         GENDER, edu, edu_a1, 
         income_r, income_a1,
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding) %>% 
  as.data.frame()

# Proportion of missing values 
prop_missing <- cbind(
  "#" = table(ifelse(complete.cases(dat_impcheck), 'complete', 'incompl.')),
  '%' = round(100 * sort(table(complete.cases(dat_impcheck)))/nrow(dat_impcheck) , 2)
)

# Proportion of missing values per variable
prop_missing_pervar <- as.data.frame(cbind(
  "# NA" = sort(colSums(is.na(dat_impcheck))),
  "% NA" = round(sort(colMeans(is.na(dat_impcheck))) * 100, 2)))

kable(prop_missing) %>%
  kable_styling(bootstrap_options = c("striped", "hover", 'responsive'), full_width = F )

prop_missing_pervar %>% 
  filter(`# NA` != 0) %>% 
  tibble::rownames_to_column(var = 'Variables') %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", 'condensed'), full_width = T )

# write.table(tbl_missing, file = 'proportion of missings_1y.txt', sep = ',', quote = F, row.names = F)

```


## Covariates imputation

### 1. Missing patterns

```{r missing patterns 13y,  fig.width = 10 , fig.height = 30}

md_pattern(dat_impcheck, pattern = T)$plot

```


------
## DATA 9 and 13 years
## Description: unimputed data

### 1. Distribution

The plots below shows the distributions of exposures, outcomes and covariates.

```{r,  fig.width = 10 , fig.height = 10}
dat_num <- dat_9_13y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy, AGE_M_v2, 
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09, 
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
        glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         age_child_mri_f09, age_child_mri_f13, DietScore_8y, gsi, BMI_0) %>% 
  as.data.frame()

par(mfrow = c(4,4))

plot_all(dat_num)

```


```{r,  fig.width = 10 , fig.height = 10}
dat_cat <- dat_9_13y %>% 
  select(GENDER, edu, income_r, ethm, ethc, SMOKE_ALL, drinkm, folic_acid, breastfeeding) %>% 
  as.data.frame()

par(mfrow = c(3,3))

plot_all(dat_cat)


```

### 2. Correlation matrix
```{r,  fig.width = 15 , fig.height = 15}
dat_9_13y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy,
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09,
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f13, 
         GENDER, edu, edu_a1, 
         income_r, income_a1,
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding) %>% 
  mutate_if(is.factor, as.numeric) %>% 
  cor(use = 'pairwise.complete.obs', method = 'spearman') %>% 
  ggcorrplot(hc.order = TRUE, 
             type = "lower",
             lab = TRUE)

```

### 3. Missing percentage
```{r}
dat_impcheck <- dat_9_13y %>% 
  select(DietScore_pregnancy, Energy_kcal_pregnancy,
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09,
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f13, 
         GENDER, edu, edu_a1, 
         income_r, income_a1,
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding) %>% 
  as.data.frame()

# Proportion of missing values 
prop_missing <- cbind(
  "#" = table(ifelse(complete.cases(dat_impcheck), 'complete', 'incompl.')),
  '%' = round(100 * sort(table(complete.cases(dat_impcheck)))/nrow(dat_impcheck) , 2)
)

# Proportion of missing values per variable
prop_missing_pervar <- as.data.frame(cbind(
  "# NA" = sort(colSums(is.na(dat_impcheck))),
  "% NA" = round(sort(colMeans(is.na(dat_impcheck))) * 100, 2)))

kable(prop_missing) %>%
  kable_styling(bootstrap_options = c("striped", "hover", 'responsive'), full_width = F )

prop_missing_pervar %>% 
  filter(`# NA` != 0) %>% 
  tibble::rownames_to_column(var = 'Variables') %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", 'condensed'), full_width = T )

# write.table(tbl_missing, file = 'proportion of missings_1y.txt', sep = ',', quote = F, row.names = F)

```


## Covariates imputation

### 1. Missing patterns

```{r missing patterns 9en13y,  fig.width = 10 , fig.height = 30}

md_pattern(dat_impcheck, pattern = T)$plot

```

