---
title: 'Imputation: maternal diet and brain'
author: "Yuchan Mou"
date: "01/Feb/2024"
output: html_document
---

```{r global settings, include=FALSE, message=FALSE}
# Global settings

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.table.format = "html")
```

```{r set working directory, include=FALSE, message=FALSE}
# WD settings
getwd()

```

````{r installpck, message = FALSE}
# install packages

.libPaths()
Sys.setenv(JAGS_HOME = r"(V:\HomeDir\044073(J. Mou)\program\JAGS)")

packages <- c('haven', 'tidyverse', 'sjmisc', 'ggcorrplot', 'kableExtra', 'mice', 'JointAI')
invisible(lapply(packages, library, character.only = T))
```


------
```{r load data}
load(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\DataManipulation_maternaldietBrain_2022-11-30.RData)')

```

------
## DATA 9 years
## Imputation

Covariates to impute are:

1) Categorical variables:
- Maternal education: edu;
    auxillary var: edu_a1

- household income: income_r;
    auxillary var: income_a1

- maternal child ethnicity: ethm;
    auxillary var: ethc

- smoking during pregnancy; SMOKE_ALL; 

- alcohol use during pregnancy: drinkm;

- folic acid use: folic_acid;

2) Numeric variables:
- child diet quality@8: DietScore_8y;

- psychopathological symptoms:gsi;

- maternal prepregnancy BMI: BMI_0;(not incl)
    

```{r imputation dataset}
dat_imp_9y <- dat_9y %>% 
  select(IDC, IDM, dti_has_nii_f09, dti_man_qc_f09, dti_auto_qc_f09, dti_overall_qc_f09,
         DietScore_pregnancy, Energy_kcal_pregnancy, 
         
         AGE_M_v2, parity, TWIN, BMI_0, WEIGHT, GESTBIR, # birth cov  
         
         CerebralWhiteMatterVol_f09, CortexVol_f09, SubCortGrayVol_f09, SupraTentorialVol_f09, genr_tbv_f09, eTIV_f09,
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, 
         glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09,
         
         CerebralWhiteMatterVol_f13, CortexVol_f13, SubCortGrayVol_f13, SupraTentorialVol_f13, genr_tbv_f13, eTIV_f13,
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, 
         glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f09, 
         GENDER, edu, edu_a1, 
         income_r, income_a1,
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding, 
         PE_total, DIAB_GRA, HYPERTENSIE_QUEST, HYPERTEN_LVR, # pregnancy complications
         visitF13, AGECHILD13, WISC13_FSIQ, WISC13_Voc_Tscore, WISC13_MR_Tscore, WISC13_DS_Tscore, WISC13_CD_Tscore,
         qdec_has_lgi_f09)
```

```{r iteration 0}
imp0 <- mice(dat_imp_9y, defaultMethod = c('pmm', "logreg", "polyreg", 'polr'), maxit = 0)
# warning dti filter indexes. 
```


```{r adjust imputation methods}
meth <- imp0$method

meth[c("dti_has_nii_f09", "dti_man_qc_f09", "dti_auto_qc_f09", "dti_overall_qc_f09", 
       "glb_mean_FA_f09", "cigulum_mean_FA_f09", "uncinate_mean_FA_f09", 
       "glb_mean_MD_f09", "cigulum_mean_MD_f09", "uncinate_mean_MD_f09",
       "prefrontal_vol_f13", "hippocampus_vol_f13", "amygdala_vol_f13", 
       "glb_mean_FA_f13", "cigulum_mean_FA_f13", "uncinate_mean_FA_f13",
       "glb_mean_MD_f13", "cigulum_mean_MD_f13", "uncinate_mean_MD_f13",
       "PE_total", "DIAB_GRA", "HYPERTENSIE_QUEST", "HYPERTEN_LVR",
       "visitF13", "AGECHILD13")] <- rep("", 25)

meth
```


```{r adjust predictor matrix}
pred <- imp0$predictorMatrix

pred=quickpred(dat_imp_9y, minpuc = 0.4, mincor=0.2, 
               include = c("ethm", "income_r", "genr_tbv_f09", "DietScore_pregnancy"))
pred[c("IDC", "IDM", 
         "dti_has_nii_f09", "dti_man_qc_f09", "dti_auto_qc_f09", "dti_overall_qc_f09", 
       
         "glb_mean_FA_f09", "cigulum_mean_FA_f09", "uncinate_mean_FA_f09", 
         "glb_mean_MD_f09", "cigulum_mean_MD_f09", "uncinate_mean_MD_f09", 
               
         
         "CerebralWhiteMatterVol_f13", "CortexVol_f13", "SubCortGrayVol_f13", "SupraTentorialVol_f13", "genr_tbv_f13", "eTIV_f13",
         "prefrontal_vol_f13", "hippocampus_vol_f13", "amygdala_vol_f13", 
         "glb_mean_FA_f13", "cigulum_mean_FA_f13", "uncinate_mean_FA_f13", 
         "glb_mean_MD_f13", "cigulum_mean_MD_f13", "uncinate_mean_MD_f13",
         "PE_total", "DIAB_GRA", "HYPERTENSIE_QUEST", "HYPERTEN_LVR", 
         "visitF13", "AGECHILD13", 
         "WISC13_Voc_Tscore", "WISC13_MR_Tscore", "WISC13_DS_Tscore", "WISC13_CD_Tscore"), ] <- 0

pred[, c("IDC", "IDM", 
         "dti_has_nii_f09", "dti_man_qc_f09", "dti_auto_qc_f09", "dti_overall_qc_f09", 
         
         "glb_mean_FA_f09", "cigulum_mean_FA_f09", "uncinate_mean_FA_f09", 
         "glb_mean_MD_f09", "cigulum_mean_MD_f09", "uncinate_mean_MD_f09", 
         
          "CerebralWhiteMatterVol_f13", "CortexVol_f13", "SubCortGrayVol_f13", "SupraTentorialVol_f13", "genr_tbv_f13", "eTIV_f13",
         "prefrontal_vol_f13", "hippocampus_vol_f13", "amygdala_vol_f13", 
         "glb_mean_FA_f13", "cigulum_mean_FA_f13", "uncinate_mean_FA_f13", 
         "glb_mean_MD_f13", "cigulum_mean_MD_f13", "uncinate_mean_MD_f13",
         "PE_total", "DIAB_GRA", "HYPERTENSIE_QUEST", "HYPERTEN_LVR", 
         "visitF13", "AGECHILD13", 
         "WISC13_Voc_Tscore", "WISC13_MR_Tscore", "WISC13_DS_Tscore", "WISC13_CD_Tscore")] <- 0

table(apply(pred,1,sum))
mean(apply(pred,1,sum))
names(dat_imp_9y)[which(apply(pred,1,sum) == 0)]

```

```{r proportion plot for convergence check}
propplot <- function(x, formula, facet = "wrap", ...) {
  library(ggplot2)
  
  
  cd <- data.frame(mice::complete(x, "long", include = TRUE))
  cd$.imp <- factor(cd$.imp)
  
  r <- as.data.frame(is.na(x$data))
  
  impcat <- x$meth != "" & sapply(x$data, is.factor)
  vnames <- names(impcat)[impcat]
  
  if (missing(formula)) {
    formula <- as.formula(paste(paste(vnames, collapse = "+",
                                      sep = ""), "~1", sep = ""))
  }
  
  tmsx <- terms(formula[-3], data = x$data)
  xnames <- attr(tmsx, "term.labels")
  xnames <- xnames[xnames %in% vnames]
  
  if (paste(formula[3]) != "1") {
    wvars <- gsub("[[:space:]]*\\|[[:print:]]*", "", paste(formula)[3])
    # wvars <- all.vars(as.formula(paste("~", wvars)))
    wvars <- attr(terms(as.formula(paste("~", wvars))), "term.labels")
    if (grepl("\\|", formula[3])) {
      svars <- gsub("[[:print:]]*\\|[[:space:]]*", "", paste(formula)[3])
      svars <- all.vars(as.formula(paste("~", svars)))
    } else {
      svars <- ".imp"
    }
  } else {
    wvars <- NULL
    svars <- ".imp"
  }
  
  for (i in seq_along(xnames)) {
    xvar <- xnames[i]
    select <- cd$.imp != 0 & !r[, xvar]
    cd[select, xvar] <- NA
  }
  
  
  for (i in which(!wvars %in% names(cd))) {
    cd[, wvars[i]] <- with(cd, eval(parse(text = wvars[i])))
  }
  
  meltDF <- reshape2::melt(cd[, c(wvars, svars, xnames)], id.vars = c(wvars, svars))
  meltDF <- meltDF[!is.na(meltDF$value), ]
  
  
  wvars <- if (!is.null(wvars)) paste0("`", wvars, "`")
  
  a <- plyr::ddply(meltDF, c(wvars, svars, "variable", "value"), plyr::summarize,
                   count = length(value))
  b <- plyr::ddply(meltDF, c(wvars, svars, "variable"), plyr::summarize,
                   tot = length(value))
  mdf <- merge(a,b)
  mdf$prop <- mdf$count / mdf$tot
  
  plotDF <- merge(unique(meltDF), mdf)
  plotDF$value <- factor(plotDF$value,
                         levels = unique(unlist(lapply(x$data[, xnames], levels))),
                         ordered = T)
  
  p <- ggplot(plotDF, aes(x = value, fill = get(svars), y = prop)) +
    geom_bar(position = "dodge", stat = "identity") +
    theme(legend.position = "bottom", ...) +
    ylab("proportion") +
    scale_fill_manual(name = "",
                      values = c("black",
                                 colorRampPalette(
                                   RColorBrewer::brewer.pal(9, "Blues"))(x$m + 3)[1:x$m + 3])) +
    guides(fill = guide_legend(nrow = 1))
  
  if (facet == "wrap")
    if (length(xnames) > 1) {
      print(p + facet_wrap(c("variable", wvars), scales = "free"))
    } else {
      if (is.null(wvars)) {
        print(p)
      } else {
        print(p + facet_wrap(wvars, scales = "free"))
      }
    }
  
  if (facet == "grid")
    if (!is.null(wvars)) {
      print(p + facet_grid(paste(paste(wvars, collapse = "+"), "~ variable"),
                           scales = "free"))
    }
}
```


```{r iteration 20 dataset 5}
imp1 <- mice(dat_imp_9y, method = meth, predictorMatrix = pred, seed = 2020, maxit = 10, m = 5)

# convergence
# traceplot
plot(imp1, layout = c(6,5))


# density plots
densityplot(imp1, ~gsi)
densityplot(imp1, ~DietScore_8y)

#~"DietScore_pregnancy, "gsi", "SDHCG3", "sdsbmiforage9child")
propplot(imp1, strip.text = element_text(size = 14))

```


```{r iteration 50 dataset 10}
imp_9y <- mice(dat_imp_9y, method = meth, predictorMatrix = pred, seed = 2021, maxit = 50, m = 10)

# convergence
# traceplot
plot(imp_9y, layout = c(6,5))

# density plots
densityplot(imp_9y, ~gsi)
densityplot(imp_9y, ~DietScore_8y)
densityplot(imp1, ~WEIGHT)

#~"DietScore_pregnancy, "gsi", "SDHCG3", "sdsbmiforage9child")
propplot(imp_9y, strip.text = element_text(size = 14))
```

------
## DATA 13 years
## Imputation

Covariates to impute are:

1) Categorical variables:
- Maternal education: edu;
    auxillary var: edu_a1

- household income: income_r;
    auxillary var: income_a1

- maternal child ethnicity: ethm;
    auxillary var: ethc

- smoking during pregnancy; SMOKE_ALL; 

- alcohol use during pregnancy: drinkm;

- folic acid use: folic_acid;

2) Numeric variables:
- child diet quality@8: DietScore_8y;

- psychopathological symptoms:gsi;

- maternal prepregnancy BMI: BMI_0;(not incl)
    

```{r imputation dataset}
dat_imp_13y <- dat_13y %>% 
  select(IDC, IDM, 
         dti_has_nii_f13, dti_man_qc_f13, dti_auto_qc_f13, dti_overall_qc_f13,
         DietScore_pregnancy, Energy_kcal_pregnancy, 
         
         AGE_M_v2, parity, TWIN, BMI_0, WEIGHT, GESTBIR, # birth cov  

         CerebralWhiteMatterVol_f09, CortexVol_f09, SubCortGrayVol_f09, SupraTentorialVol_f09, genr_tbv_f09, eTIV_f09,
         prefrontal_vol_f09, hippocampus_vol_f09, amygdala_vol_f09, 
         glb_mean_FA_f09, cigulum_mean_FA_f09, uncinate_mean_FA_f09,
         glb_mean_MD_f09, cigulum_mean_MD_f09, uncinate_mean_MD_f09,
         
         CerebralWhiteMatterVol_f13, CortexVol_f13, SubCortGrayVol_f13, SupraTentorialVol_f13, genr_tbv_f13, eTIV_f13,
         prefrontal_vol_f13, hippocampus_vol_f13, amygdala_vol_f13, 
         glb_mean_FA_f13, cigulum_mean_FA_f13, uncinate_mean_FA_f13,
         glb_mean_MD_f13, cigulum_mean_MD_f13, uncinate_mean_MD_f13,
         
         AGE_M_v2, DietScore_8y, gsi, 
         age_child_mri_f13, 
         GENDER, edu, edu_a1, 
         income_r, income_a1, 
         ethm, ethc, drinkm, SMOKE_ALL, folic_acid, breastfeeding,
         PE_total, DIAB_GRA, HYPERTENSIE_QUEST, HYPERTEN_LVR, # pregnancy complications
         visitF13, AGECHILD13, WISC13_FSIQ, WISC13_Voc_Tscore, WISC13_MR_Tscore, WISC13_DS_Tscore, WISC13_CD_Tscore)
```

```{r iteration 0}
imp0 <- mice(dat_imp_13y, defaultMethod = c('pmm', "logreg", "polyreg", 'polr'), maxit = 0)

```


```{r adjust imputation methods}
meth <- imp0$method

meth[c("dti_has_nii_f13", "dti_man_qc_f13", "dti_auto_qc_f13", "dti_overall_qc_f13", 
       
        "CerebralWhiteMatterVol_f09", "CortexVol_f09", "SubCortGrayVol_f09", "SupraTentorialVol_f09", "genr_tbv_f09", "eTIV_f09",
       "prefrontal_vol_f09", "hippocampus_vol_f09", "amygdala_vol_f09",
       "glb_mean_FA_f09", "cigulum_mean_FA_f09", "uncinate_mean_FA_f09", 
       "glb_mean_MD_f09", "cigulum_mean_MD_f09", "uncinate_mean_MD_f09",
        
       "CerebralWhiteMatterVol_f13", "CortexVol_f13", "SubCortGrayVol_f13", "SupraTentorialVol_f13", "genr_tbv_f13", "eTIV_f13",
       "prefrontal_vol_f13", "hippocampus_vol_f13", "amygdala_vol_f13", 
       "glb_mean_FA_f13", "cigulum_mean_FA_f13", "uncinate_mean_FA_f13",
       "glb_mean_MD_f13", "cigulum_mean_MD_f13", "uncinate_mean_MD_f13",
       
       "PE_total", "DIAB_GRA", "HYPERTENSIE_QUEST", "HYPERTEN_LVR",
       "visitF13", "AGECHILD13")] <- rep("", 40)

meth
```


```{r adjust predictor matrix}
pred <- imp0$predictorMatrix

pred=quickpred(dat_imp_13y, minpuc = 0.4, mincor=0.2, 
               include = c("ethm", "income_r", "genr_tbv_f13", "DietScore_pregnancy"))

pred[, c("IDC", "IDM", 
         "dti_has_nii_f13", "dti_man_qc_f13", "dti_auto_qc_f13", "dti_overall_qc_f13",
         
         "CerebralWhiteMatterVol_f09", "CortexVol_f09", "SubCortGrayVol_f09", "SupraTentorialVol_f09", "genr_tbv_f09", "eTIV_f09",
        "prefrontal_vol_f09", "hippocampus_vol_f09", "amygdala_vol_f09",
        "glb_mean_FA_f09", "cigulum_mean_FA_f09", "uncinate_mean_FA_f09", 
        "glb_mean_MD_f09", "cigulum_mean_MD_f09", "uncinate_mean_MD_f09",
         
         "glb_mean_FA_f13", "cigulum_mean_FA_f13", "uncinate_mean_FA_f13", 
         "glb_mean_MD_f13", "cigulum_mean_MD_f13", "uncinate_mean_MD_f13",
        
         "PE_total", "DIAB_GRA", "HYPERTENSIE_QUEST", "HYPERTEN_LVR", 
         "visitF13", "AGECHILD13", 
         "WISC13_Voc_Tscore", "WISC13_MR_Tscore", "WISC13_DS_Tscore", "WISC13_CD_Tscore")] <- 0

pred[c("IDC", "IDM", 
         "dti_has_nii_f13", "dti_man_qc_f13", "dti_auto_qc_f13", "dti_overall_qc_f13",
         
         "CerebralWhiteMatterVol_f09", "CortexVol_f09", "SubCortGrayVol_f09", "SupraTentorialVol_f09", "genr_tbv_f09", "eTIV_f09",
        "prefrontal_vol_f09", "hippocampus_vol_f09", "amygdala_vol_f09",
        "glb_mean_FA_f09", "cigulum_mean_FA_f09", "uncinate_mean_FA_f09", 
        "glb_mean_MD_f09", "cigulum_mean_MD_f09", "uncinate_mean_MD_f09",
         
         "glb_mean_FA_f13", "cigulum_mean_FA_f13", "uncinate_mean_FA_f13", 
         "glb_mean_MD_f13", "cigulum_mean_MD_f13", "uncinate_mean_MD_f13",
       
         "PE_total", "DIAB_GRA", "HYPERTENSIE_QUEST", "HYPERTEN_LVR", 
         "visitF13", "AGECHILD13", 
         "WISC13_Voc_Tscore", "WISC13_MR_Tscore", "WISC13_DS_Tscore", "WISC13_CD_Tscore"), ] <- 0

table(apply(pred,1,sum))
mean(apply(pred,1,sum))
names(dat_imp_13y)[which(apply(pred,1,sum) == 0)]

```

```{r iteration 20 dataset 5}
imp1 <- mice(dat_imp_13y, method = meth, predictorMatrix = pred, seed = 2020, maxit = 10, m = 5)

# convergence
# traceplot
plot(imp1, layout = c(6,5))


# density plots
densityplot(imp1, ~gsi)
densityplot(imp1, ~DietScore_8y)

#~"DietScore_pregnancy, "gsi", "SDHCG3", "sdsbmiforage9child")
propplot(imp1, strip.text = element_text(size = 14))

```


```{r iteration 50 dataset 10}
imp_13y <- mice(dat_imp_13y, method = meth, predictorMatrix = pred, seed = 2021, maxit = 50, m = 10)

# convergence
# traceplot
plot(imp_13y, layout = c(6,5))


# density plots
densityplot(imp_13y, ~gsi)
densityplot(imp_13y, ~DietScore_8y)
densityplot(imp_13y, ~WEIGHT)

#~"DietScore_pregnancy, "gsi", "SDHCG3", "sdsbmiforage9child")
propplot(imp_13y, strip.text = element_text(size = 14))


```


# Add variables
## add raw WISC subscores
```{r}
load(r"(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\imp_mdq_brain_2022-11-30.RData)")

fsiq_13y <- haven::read_spss(r"(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\CHILDWISC13_16082021.sav)") %>% 
  arrange(IDC) %>% 
  select(IDC, visitF13, AGECHILD13, WISC13_Voc_Raw, WISC13_MR_Raw, WISC13_DS_Raw, WISC13_CD_Raw)

bcore <- readRDS(r'(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\genr_mri_core_data_20220311.rds)') %>% 
  arrange(idc) %>% 
  rename(IDC = idc) %>% 
  select(IDC, folders_f09, folders_f13)
```


```{r}
#Data Manipulation ------------------------

# dti data
imp_9y_dti <- mice::filter(imp_9y, dti_has_nii_f09 == "yes" & dti_man_qc_f09 != "unusable" )

# maternal diet quality tertiles
imp_9y_quartile <- imp_9y %>% 
  mice::complete(action = "long", include = T) 

# 1st quartile: 6.796, median 7.833, 3rd quartile: 8.875

imp_9y_quartile <- mutate(imp_9y_quartile, mdq_quartile = ntile(DietScore_pregnancy, 4))

imp_9y_quartile$mdq_quartile <- factor(imp_9y_quartile$mdq_quartile,levels = c(1,2,3, 4), 
                                      labels = c("Q1", "Q2", "Q3", "Q4"),
                                      ordered = F)  
imp_9y_quartile$mdq_quartile <- relevel(imp_9y_quartile$mdq_quartile, ref = "Q1")

table(imp_9y_quartile$mdq_quartile, useNA = "always")

# child diet quality tertile
# 1st qu 3.7, 3rd qu 5.4n

imp_9y_quartile <- mutate(imp_9y_quartile, cdq_tertile = ntile(DietScore_8y, 3))

imp_9y_quartile$cdq_tertile <- factor(imp_9y_quartile$cdq_tertile,levels = c(1,2,3), 
                                      labels = c("low", "mid", "high"),
                                      ordered = F)

table(imp_9y_quartile$cdq_tertile, useNA = "always")


# add IQ sub scores
imp_9y_quartile <- imp_9y_quartile %>% 
  arrange(IDC) %>% 
  left_join(fsiq_13y, by = 'IDC')


# add folder variable - as id indicator in qdecr
imp_9y_quartile <- imp_9y_quartile %>% 
  arrange(IDC) %>% 
  left_join(bcore, by = 'IDC')  

#back to mids
imp_9y <- as.mids(imp_9y_quartile, .id = ".id")


## dti dataset


# maternal diet quality tertiles
imp_9y_dti_quartile <- imp_9y_dti %>% 
  mice::complete(action = "long", include = T) 

# 1st quartile: 6.7, 3rd quartile: 8.9

imp_9y_dti_quartile <- mutate(imp_9y_dti_quartile, mdq_quartile = ntile(DietScore_pregnancy, 4))

imp_9y_dti_quartile$mdq_quartile <- factor(imp_9y_dti_quartile$mdq_quartile,levels = c(1,2,3,4), 
                                     labels = c("Q1", "Q2", "Q3", "Q4"),
                                     ordered = F)  
imp_9y_dti_quartile$mdq_quartile <- relevel(imp_9y_dti_quartile$mdq_quartile, ref = "Q1")

table(imp_9y_dti_quartile$mdq_quartile, useNA = "always")

# child diet quality tertile
# 1st qu 3.7, 3rd qu 5.4

imp_9y_dti_quartile <- mutate(imp_9y_dti_quartile, cdq_tertile = ntile(DietScore_8y, 3))

imp_9y_dti_quartile$cdq_tertile <- factor(imp_9y_dti_quartile$cdq_tertile,levels = c(1,2,3), 
                                     labels = c("low", "mid", "high"),
                                     ordered = F)

table(imp_9y_dti_quartile$cdq_tertile, useNA = "always")


#back to mids
imp_9y_dti <- as.mids(imp_9y_dti_quartile, .id = ".id")
```


```{r}
#Data Manipulation ------------------------

# dti data
imp_13y_dti <- mice::filter(imp_13y, dti_has_nii_f13 == "yes" & dti_man_qc_f13 != "unusable" )

# maternal diet quality tertiles
imp_13y_quartile <- imp_13y %>% 
  mice::complete(action = "long", include = T) 

# 1st quartile: 6.796, median 7.833, 3rd quartile: 8.875

imp_13y_quartile <- mutate(imp_13y_quartile, mdq_quartile = ntile(DietScore_pregnancy, 4))

imp_13y_quartile$mdq_quartile <- factor(imp_13y_quartile$mdq_quartile,levels = c(1,2,3, 4), 
                                        labels = c("Q1", "Q2", "Q3", "Q4"),
                                        ordered = F)  
imp_13y_quartile$mdq_quartile <- relevel(imp_13y_quartile$mdq_quartile, ref = "Q1")

table(imp_13y_quartile$mdq_quartile, useNA = "always")

# child diet quality tertile
# 1st qu 3.7, 3rd qu 5.4n

imp_13y_quartile <- mutate(imp_13y_quartile, cdq_tertile = ntile(DietScore_8y, 3))

imp_13y_quartile$cdq_tertile <- factor(imp_13y_quartile$cdq_tertile,levels = c(1,2,3), 
                                       labels = c("low", "mid", "high"),
                                       ordered = F)

table(imp_13y_quartile$cdq_tertile, useNA = "always")

# add IQ sub scores
imp_13y_quartile <- imp_13y_quartile %>% 
  arrange(IDC) %>% 
  left_join(fsiq_13y, by = 'IDC')


# add folder variable - as id indicator in qdecr
imp_13y_quartile <- imp_13y_quartile %>% 
  arrange(IDC) %>% 
  left_join(bcore, by = 'IDC')  


#back to mids
imp_13y <- as.mids(imp_13y_quartile, .id = ".id")


## dti dataset


# maternal diet quality tertiles
imp_13y_dti_quartile <- imp_13y_dti %>% 
  mice::complete(action = "long", include = T) 

# 1st quartile: 6.7, 3rd quartile: 8.9

imp_13y_dti_quartile <- mutate(imp_13y_dti_quartile, mdq_quartile = ntile(DietScore_pregnancy, 4))

imp_13y_dti_quartile$mdq_quartile <- factor(imp_13y_dti_quartile$mdq_quartile,levels = c(1,2,3,4), 
                                            labels = c("Q1", "Q2", "Q3", "Q4"),
                                            ordered = F)  
imp_13y_dti_quartile$mdq_quartile <- relevel(imp_13y_dti_quartile$mdq_quartile, ref = "Q1")

table(imp_13y_dti_quartile$mdq_quartile, useNA = "always")

# child diet quality tertile
# 1st qu 3.7, 3rd qu 5.4

imp_13y_dti_quartile <- mutate(imp_13y_dti_quartile, cdq_tertile = ntile(DietScore_8y, 3))

imp_13y_dti_quartile$cdq_tertile <- factor(imp_13y_dti_quartile$cdq_tertile,levels = c(1,2,3), 
                                           labels = c("low", "mid", "high"),
                                           ordered = F)

table(imp_13y_dti_quartile$cdq_tertile, useNA = "always")


#back to mids
imp_13y_dti <- as.mids(imp_13y_dti_quartile, .id = ".id")
```


Final dataset after imputation is imp_9y and imp_13y.


Save as RDS file



```{r}

analysis_date = Sys.Date()

save(imp_9y, file = paste(r"(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\imp_mdq_brain_9y_)",
      analysis_date, ".RData", sep = ""))

save(imp_13y, file = paste(r"(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\imp_mdq_brain_13y_)",
      analysis_date, ".RData", sep = ""))

save(imp_9y, imp_13y, file = paste(r"(V:\HomeDir\044073(J. Mou)\Projects\4 Maternal diet and brain morphology\data\imp_mdq_brain_)", analysis_date, ".RData", sep = ""))
```

